Lyman-alpha spectra from rotating outflows

### 2018-03-27 paper/paper.tex
new references

These kinematic maps show that LAEs have two  overimposed velocity
patterns: coherent rotation plus and a random motions.
Depending on the system under study, one of the two features can be
more important, or they can be found in equal measure.
These kinematic features play an important role in shaping the \lya.
Models of the \lya have also shown that there is clear signal for
outflowing gas most probably produced by star formation activity.
The characteristic signal for these outflows is the asymmetrical \lya
linea with a more pronounced read peak.
The effects of either outflows or rotation on the \lya line have been
already explored.

@@

We found that in the case of solid body rotation these effects can be quantitatively
explained by a Doppler boost, where the Doppler factor can be computed as the
 producet of quantities at the surface of last scattering, namely $\vec{v}_{\rm rot}\cdot \hat{k}$,
 where $\vec{v}_{\rm rot}$ is the velocity due to rotation and $\hat{k}$ is the direction
 of the photon's propagation.
These conclusions trenghten the evidence reported by \cite{Garavito14} when they modeled
the effect of pure rotation on \lya\ spectra.

>>>>

Dwarf galaxies show signs of coherent rotation
\citep{2009A&A...493..871S}.
Some of them are expected to have high neutral gas contents and thus
show \lya\ emission with rotation imprints
\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
Some Compact Dwarf Galaxies actually show gas kinematics that range
from pure rotation to high velocity dispersion without a clear
rotation pattern.
\citep{2015A&A...577A..21C,2017arXiv170809407C,2017A&A...600A.125C}

The case for outflows as traced in the \lya\ line is much bettwe
stablished.
There are abundance cases \lya line profiles with a single peak
redwards from the line's center. 
In other cases there is a double peak but the peak on the red side is
stronger \citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.  
This has been readily explained as the consequence of multiple \lya photon
scatterings through an homogeneous outflowing shell of neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

@@

We found that in the case of solid body rotation these effects can be quantitatively
explained by a Doppler boost, where the Doppler factor can be computed as the
 producet of quantities at the surface of last scattering, namely $\vec{v}_{\rm rot}\cdot \hat{k}$,
 where $\vec{v}_{\rm rot}$ is the velocity due to rotation and $\hat{k}$ is the direction
of the photon's propagation.
These conclusions, specially the confirmation of the Doppler boost as
a good proxy for solid body rotation, strenghten the evidence reported
by \cite{Garavito14} when they modeled the effect of pure rotation on
\lya\ spectra.

As an application to observational data we find that recent results
that take the spectra of two different sides of a galaxy can detect 
the effect of approaching/receeding gas motions. 
However, the distances between the peaks of these two spectra 
correspond to $\approx v_{\rm rot} \cos\theta$ due to the weights of the
emitting regions.
